### 2014-02-26 paper/Rcode/ashsim.rmd
rearranged simulations for paper some

```{r}
  require("ashr")
  ```{r}
opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
require(ashr)
require(qvalue)
require(fdrtool)
require(mixfdr)
require(locfdr)
require(ggplot2)
#simulate from mixture of normals
#compare normal, uniform and half-uniform mixtures
#bsd gives standard deviation of beta
ashsim=function(mixmean,mixsd,mixpi,bsd=1,seedval = 100,nsamp=1000,niter=50){  
  set.seed(seedval)  
  beta =list()
  betahatsd=list()
  betahat = list()
  betahat.ash.n = list()
  betahat.ash.u = list()
  betahat.ash.hu = list()
  k = length(mixmean)
  for(i in 1:niter){
    comp = sample(1:k,nsamp,prob=mixpi,replace=TRUE)
    sd = mixsd[comp]
    mean = mixmean[comp]
    beta[[i]] = rnorm(nsamp,mean,sd)
    betahatsd[[i]] = bsd
    betahat[[i]] = beta[[i]]+rnorm(nsamp,0,betahatsd[[i]])
    betahat.ash.n[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="normal",method="shrink")
    betahat.ash.u[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="uniform", method="shrink")
    betahat.ash.hu[[i]] = ash(betahat[[i]],betahatsd[[i]],mixcompdist="halfuniform", method="shrink")

  }
  return(list(beta =beta,
  betahatsd=betahatsd,
  betahat = betahat,
  betahat.ash.n = betahat.ash.n,
  betahat.ash.u = betahat.ash.u,
  betahat.ash.hu = betahat.ash.hu))
}    
```

```{r}
sim1= ashsim(c(0,0,0),c(1,1,2),c(1/3,1/3,1/3),niter=20,nsamp=1000)
sim2= ashsim(c(-1.5,-1,-0.5,0,0.5,1,1.5),rep(0.5,7),rep(1/7,7),niter=20,nsamp=1000)
sim3= ashsim(c(-2,-1,0,1),c(2,1.5,1,1),c(1/4,1/4,1/3,1/6),niter=20,nsamp=1000)

@@

```{r}
len=length(sim1$beta[[1]])
res = data.frame(beta = c(sim1$beta[[1]],sim2$beta[[1]],sim3$beta[[1]]), Scenario = c(rep(1,len),rep(2,len),rep(3,len)))
x = seq(-8,6,length=100)
y = dnorm(x)
dens1 = data.frame(x=x,dn= t(density(sim1$betahat.ash.n[[1]],x)$y),du=t(density(sim1$betahat.ash.u[[1]],x)$y),dhu=t(density(sim1$betahat.ash.hu[[1]],x)$y), Scenario=1)
dens2 = data.frame(x=x,dn = t(density(sim2$betahat.ash.n[[1]],x)$y),du=t(density(sim2$betahat.ash.u[[1]],x)$y), dhu=t(density(sim2$betahat.ash.hu[[1]],x)$y), Scenario=2)
dens3 = data.frame(x=x,dn = t(density(sim3$betahat.ash.n[[1]],x)$y),du=t(density(sim3$betahat.ash.u[[1]],x)$y), dhu=t(density(sim3$betahat.ash.hu[[1]],x)$y), Scenario=3)
dens=rbind(dens1,dens2,dens3)
ggplot(res, aes(x=beta)) + facet_grid(.~Scenario)  +
    geom_histogram(aes(y=..density..),binwidth=0.5,
                   colour="black", fill="white") +  
  geom_line(data = dens, aes(x,du,color='Uniform'),size=1.2,alpha=0.8) +
  geom_line(data = dens, aes(x,dhu,color='Half Uniform'),size=1.2,alpha=0.8) +
    geom_line(data = dens, aes(x,dn,color='Normal'),size=1.2,alpha=0.8) +
  scale_colour_manual(name = 'Method', values = c('red', 'blue','green'))

@@

```{r}
  rmse = function(x,y){sqrt(mean((x-y)^2))}
  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
  plot_rmse = function(sims,inczero=FALSE,incbetahat=FALSE){
    res=list()
    res2=list()
    for(i in 1:length(sims)){
      err.ash.n= mapply(get_rmse.ash,sims[[i]]$betahat.ash.n,sims[[i]]$beta)
      err.ash.u = mapply(get_rmse.ash,sims[[i]]$betahat.ash.u,sims[[i]]$beta)
      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$betahat.ash.hu,sims[[i]]$beta)
    #err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
      err.betahat = mapply(rmse,sims[[i]]$betahat,sims[[i]]$beta)
      err.zero = unlist(lapply(sims[[i]]$beta,rmse,y=0))
      loglik.ash.n = unlist(lapply(sims[[i]]$betahat.ash.n,get_loglik))
      loglik.ash.u = unlist(lapply(sims[[i]]$betahat.ash.u,get_loglik))
      loglik.ash.hu = unlist(lapply(sims[[i]]$betahat.ash.hu,get_loglik))
     res[[i]] = data.frame(Scenario=i,ash.normal=err.ash.n,ash.uniform=err.ash.u,ash.halfuniform=err.ash.hu)
      res2[[i]] = data.frame(Scenario=i,ash.normal=loglik.ash.n,ash.uniform=loglik.ash.u,ash.halfuniform=loglik.ash.hu)

      if(inczero){
        res[[i]]=data.frame(res,zero=err.zero)
      }
      if(incbetahat){
        res[[i]]=data.frame(res,betahat=err.betahat)
      }
    }
    require(reshape2)
    res.melt = melt(res, id.vars=c("Scenario"),variable.name="Method")
    res2.melt = melt(res2, id.vars=c("Scenario"),variable.name="Method")
    ggplot(data=res.melt,aes(Method,value)) +geom_boxplot()+         facet_grid(. ~ Scenario)
     ggplot(data=res2.melt,aes(Method,value)) +geom_boxplot()+         facet_grid(. ~ Scenario)

    p=ggplot(data=res.melt,aes(ash.halfuniform,value,colour=Method)) +geom_point(shape=16) +
         facet_grid(. ~ Scenario) +
        geom_abline(colour = "black") +
        xlab("RMSE (ash.halfuniform)") +
        ylab("RMSE (other method)")
    print(p +
          coord_equal(ratio=1))
  }
  #pdf("figures/rmse_sim1.pdf")
    plot_rmse(list(sim1,sim2,sim3))
 #               dev.off()
  #pdf("figures/rmse_sim2.pdf")
  #  plot_rmse(simres2)
  #dev.off()
```

Try plotting boxplots of rmse and fit impovement compared with most general model
```{r}
  rmse = function(x,y){sqrt(mean((x-y)^2))}
  get_rmse.ash = function(a,b){rmse(a$PosteriorMean,b)}
  get_rmse.mixfdr = function(a,b){rmse(a$effectSize,b)}
  boxplot_rmse_loglik = function(sims,inczero=FALSE,incbetahat=FALSE){
    res=list()
    res2=list()
    for(i in 1:length(sims)){
      err.ash.hu = mapply(get_rmse.ash,sims[[i]]$betahat.ash.hu,sims[[i]]$beta)
      err.ash.n= mapply(get_rmse.ash,sims[[i]]$betahat.ash.n,sims[[i]]$beta)-err.ash.hu
      err.ash.u = mapply(get_rmse.ash,sims[[i]]$betahat.ash.u,sims[[i]]$beta)-err.ash.hu
      #err.mixfdr=mapply(get_rmse.mixfdr,sims$betahat.mixfdr,sims$beta)
      loglik.ash.hu = unlist(lapply(sims[[i]]$betahat.ash.hu,get_loglik))
      loglik.ash.n = unlist(lapply(sims[[i]]$betahat.ash.n,get_loglik))-loglik.ash.hu
      loglik.ash.u = unlist(lapply(sims[[i]]$betahat.ash.u,get_loglik))-loglik.ash.hu

     res[[i]] = data.frame(Scenario=i,ash.normal.err=err.ash.n,ash.uniform.err=err.ash.u)
       res2[[i]] = data.frame(Scenario=i,ash.normal=loglik.ash.n,ash.uniform=loglik.ash.u)

    }
    require(reshape2)
    res.melt = melt(res, id.vars=c("Scenario"))

    res2.melt=melt(res2,id.vars=c("Scenario"))
    ggplot(res.melt,aes(variable,value)) + geom_boxplot() + facet_grid(.~Scenario)
    ggplot(res2.melt,aes(variable,value)) + geom_boxplot() + facet_grid(.~Scenario)
  }

  pdf("figures/boxplot_rmse_loglik.pdf")
    boxplot_rmse_loglik(list(sim1,sim2,sim3))
  dev.off()

@@

```{r}
  plot_LR=function(sims){
    hist(unlist(lapply(sims$betahat.ash.u,get_loglik))-unlist(lapply(sims$betahat.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
  }

  pdf("figures/logLR.pdf")
    plot_LR(simres1)
    plot_LR(simres2)
  dev.off()
```

```{r}
  lapply()
```

>>>>

```{r}
opts_knit$set(progress = TRUE, verbose = TRUE,root.dir="~/Documents/git/ash/paper/Rcode")
require(ashr)
require(qvalue)
require(fdrtool)
require(mixfdr)
require(locfdr)
require(ggplot2)   
```

```{r}
  load("ashsim_out.RData")
  source("plot_examples.R")
  source("plot_rmse.R")
```

```{r}
pdf("figures/simABC_egdens.pdf",width=6.5,height=2)
plot_examples(list(simA,simB,simC))
dev.off()
```

```{r}
  pdf("figures/rmse_biplot.pdf",width=6.5,height=2)
    plot_rmse(list(simA,simB,simC))
  dev.off()

#This figure not used in paper?
pdf("figures/rmse_biplot_withzerobetahat.pdf")
    plot_rmse(list(simA,simB,simC),inczero=TRUE,incbetahat=TRUE)
dev.off()

@@

```{r}

  pdf("figures/rmse_boxplot.pdf",width=6.5,height=2)
    plot_rmse_boxplot(list(simA,simB,simC))
  dev.off()

    pdf("figures/rmse_boxplot_extended.pdf",width=6.5,height=2)
    plot_rmse_boxplot(list(simA,simB,simC),TRUE,TRUE,TRUE)
  dev.off()

@@

```{r}

  pdf("figures/loglik_boxplot.pdf",width=6.5,height=2)
    plot_loglik_boxplot(list(simA,simB,simC))
  dev.off()

@@

```{r}

  pdf("figures/rmse_loglik_boxplot.pdf",width=6.5,height=5)
    plot_rmse_loglik_boxplot(list(simA,simB,simC))
  dev.off()

looking to see why mixfdr gives worse rmse, even though loglik looks ok.
Perhaps it overshrinks?
```{r}
 plot(simC$fit.ash.hu[[1]]$PosteriorMean, simC$fit.mixfdr[[1]]$effectSize)
 plot(simC$fit.ash.fdr.n[[1]]$PosteriorMean, simC$fit.mixfdr[[1]]$effectSize)
```

problem seems to be that it overestimates pi0
```{r}
 simC$fit.mixfdr[[1]]$pi
```
[1] 0.71660654 0.08298064 0.20041282

```{r}
temp=mixFdr(simC$betahat[[1]]/simC$betahatsd[[1]],noiseSD=1,theonull=TRUE,calibrate=TRUE)
temp2=mixFdr(simC$betahat[[1]]/simC$betahatsd[[1]],noiseSD=1,theonull=TRUE,J=10,plot=FALSE)
plot(ecdf(simC$beta[[1]]),xlim=c(-6,6))
lines(x,cdf.mixfdr(temp,x),col=3)
lines(cdf.ash(simC$fit.ash.n[[1]],x),col=2)
lines(cdf.ash(simC$fit.ash.hu[[1]],x),col=2,lty=2)
```

```{r}
  plot_LR=function(sims){
    hist(unlist(lapply(sims$fit.ash.u,get_loglik))-unlist(lapply(sims$fit.ash.n,get_loglik)), xlab="loglik difference", main="loglik differences for nullbiased prior vs mle",nclass=10)
  }

  pdf("figures/logLR.pdf")
    plot_LR(simA)
    plot_LR(simB)
  dev.off()
```

## Unused figures?

```{r}

pdf("figures/simABC_eg_withfit.pdf",width=6.5,height=2)
plot_examples_withfit(list(simA,simB,simC))
dev.off()
```

```{r}
pdf("figures/simABC_eg_cdf_withfit.pdf",width=6.5,height=2)
plot_examples_cdf_withfit(list(simA,simB,simC))
dev.off()
```

Experiment with mixFdr to see how penalization effects fit. 
```{r}
  betahat = simA$betahat[[1]]
  betahatsd = simA$betahatsd[[1]]
  set.seed(111) #to ensure reproducibility
  fit.mixfdr.P0 = mixFdr(betahat/betahatsd,noiseSD=1,theonull=TRUE,P=0,plot=FALSE)
  cdf.mixfdr = function(a,x){mixcdf(normalmix(a$pi,a$mu,a$sigma-1),x)}
  x = seq(-4,4,length=100)

  res = data.frame(beta=x,
                  mixfdr.P0 = t(cdf.mixfdr(fit.mixfdr.P0,x)),
                  mixfdr = t(cdf.mixfdr(simA$fit.mixfdr[[1]],x)),
                  ash.n = t(cdf.ash(simA$fit.ash.n[[1]],x)$y),
                  ash.fdr.n=t(cdf.ash(simA$fit.ash.fdr.n[[1]],x)$y))

  truth = t(cdf.ash(simA$fit.ash.true[[1]],x)$y)

  res.melt=melt(res, id.vars=c("beta"),variable.name="Method")   
  res.melt$Penalization= ifelse((res.melt$Method=="mixfdr" | res.melt$Method=="ash.fdr.n"),"Default","Minimal")
  res.melt$Methodtype=as.character(res.melt$Method)
  res.melt$Methodtype[res.melt$Method=="mixfdr.P0"]="mixfdr"
  res.melt$Methodtype[res.melt$Method=="ash.fdr.n"]="ash"
  res.melt$Methodtype[res.melt$Method=="ash.n"]="ash"

cbbPalette <- c("#56B4E9", "#009E73", "#000000","#F0E442", "#0072B2", "#E69F00",  "#D55E00", "#CC79A7")

pdf("figures/penaltycompare.pdf")
ggplot(res.melt, aes(x=beta)) +
  geom_line(aes(x=beta,y=value,color=Methodtype,linetype=Penalization,group=Method),size=1.5,alpha=0.8) +
  geom_line(data=res, aes(y=truth,color="truth"),alpha=1,size=0.5) +
  scale_colour_manual(name = 'Method', values = cbbPalette)
dev.off()
```
